---
title: "R Notebook"
author: "William Martinez"
date: '2022-05-17'
output:
  html_document:
    df_print: paged
---

# Introduction to R

```{r}
setwd("C:\\Users\\William\\OneDrive - Universidad Nacional de Colombia\\Documentos\\DOCTORATE\\Courses\\SemestreIV\\SocialSoftware\\Curso\\Modulo4\\R\\lessons")
## You can double-check that this worked by using the getwd() function. What does this do?
getwd()
pop <- read.csv("gapminder-FiveYearData.csv")

```
## Basic operations

Display dataframe

```{r}
head(pop)
```

Get basic stats

```{r}
summary(pop)
```
Select the columns continent and pop

```{r}
pop_query = dplyr::select(pop, continent,pop)
pop_query
```


## Subset Data Frames

```{r}
#install.packages("dplyr")
library(dplyr)
pop = tibble(pop)
pop
```


```{r}
select(pop, continent,pop)
```


```{r}
respop_asia = pop %>% filter(continent == "Asia")
respop_asia
```
```{r}
respop_asia = pop %>% filter(continent == "Asia")
respop_asia
```

Find population by country in 2007

```{r}
respop_2007 = pop %>% filter(year == 2007)
respop_2007
```

Find the population by country in 2007 related to countries in Asia

```{r}
respop_2007_asia = pop %>% filter(year == 2007, continent == "Asia")
respop_2007_asia
```

Challenge: Find the population by country in 2002 related to countries in America

Hint = "Americas"


```{r}

```

What countries does Europe have?

```{r}
countries_europe = pop %>% filter(continent == "Europe")
unique(countries_europe$country)
```

## Group by dataframes

Find the population in 2007 by continents

```{r}
respop_2007 = pop %>% filter(year == 2007)
pop_continent_2007 = respop_2007 %>% group_by(continent) %>% summarise(Pop = sum(pop))
pop_continent_2007
```
Find the number of countries by continent

```{r}
pop_continent_ncountries = pop %>% group_by(continent) %>% summarise(Pop = length(unique(country)))
pop_continent_ncountries
```

# graphics using ggplot

Find the number of countries by continent

```{r}
#install.packages(c("ggplot2"))
library(ggplot2)
```

create a histrogram with the life expetancy

```{r}
#install.packages(c("ggplot2"))
gg_hist <- ggplot(data = pop, aes(x = lifeExp)) + 
  # Customize bar color and number
  geom_histogram(fill = "blue", color = "black", bins = 10) + 
  # Add title
  ggtitle("Life expectancy for the gap dataset") + 
  # Change x-axis label
  xlab("Life expectancy (years)") + 
  # y-axis label
  ylab("Frequency") + 
  # Change background theme
  theme_classic() 
print(gg_hist)
```

```{r}
## Now, let's try boxplots, ggplot style:
gg_box <- ggplot(data = pop, 
                 aes(x = continent, 
                     y = lifeExp, 
                     fill = continent)) + 
  geom_boxplot() + 
  ggtitle("Boxplots for lifeExp by continent") + 
  xlab("Continent") + 
  ylab("Life expectancy (years)") +
  theme_minimal() 
print(gg_box)
```



```{r}
gg_line <- ggplot(pop, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth(method = "lm", lwd = 1, col = "red") + 
  ggtitle("gdpPercap and lifeExp") + 
  xlab("GDP per capita income (USD)") + 
  ylab("Life expectancy (years)") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 40)) +
  scale_x_log10()
print(gg_line)
```

