---
title: "World population"
author: "William Martínez"
date: "17/05/2023"
output:
  html_document:
    theme: journal
    toc: yes
    toc_depth: 4
    number_section: yes
    code_folding: hide
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
fig_caption: yes
---

# Introduction to dataframw with R

Display dataframe

```{r}
setwd("C:\\Users\\William\\OneDrive - Universidad Nacional de Colombia\\Documentos\\DOCTORATE\\Courses\\SemestreIV\\SocialSoftware\\Curso\\Modulo4\\R")
## You can double-check that this worked by using the getwd() function. What does this do?
getwd()
pop <- read.csv("gapminder-FiveYearData.csv")
knitr::kable(head(pop,5), caption = "Population worldwide")
```

## Basic operations

Get basic stats

```{r}
summary(pop)
```
Select the columns continent and pop

```{r}
pop_query = dplyr::select(pop, continent,pop)
head(pop_query,5)
```


## Subset Data Frames

```{r}
#install.packages("dplyr")
library(dplyr)
pop = tibble(pop)
head(pop,5)
```


```{r}
pop_query_continent = select(pop, continent,pop)
head(pop_query_continent,5)
```


```{r}
respop_asia = pop %>% filter(continent == "Asia")
head(respop_asia,5)
```


Find population by country in 2007

```{r}
respop_2007 = pop %>% filter(year == 2007)
head(respop_2007,5)
```

Find the population by country in 2007 related to countries in Asia

```{r}
respop_2007_asia = pop %>% filter(year == 2007, continent == "Asia")
head(respop_2007_asia,5)
```

Challenge: Find the population by country in 2002 related to countries in America

Hint = "Americas"


```{r}

```

What countries does Europe have?

```{r}
countries_europe = pop %>% filter(continent == "Europe")
unique(countries_europe$country)
```

## Group by dataframes

Find the population in 2007 by continents

```{r}
respop_2007 = pop %>% filter(year == 2007)
pop_continent_2007 = respop_2007 %>% group_by(continent) %>% summarise(Pop = sum(pop))
head(pop_continent_2007,5)
```
Find the number of countries by continent

```{r}
pop_continent_ncountries = pop %>% group_by(continent) %>% summarise(Pop = length(unique(country)))
head(pop_continent_ncountries,5)
```

# graphics using ggplot

Find the number of countries by continent

```{r ,warning = FALSE, echo=FALSE, message=FALSE}
#install.packages(c("ggplot2"))
library(ggplot2)
```

create a histogram with the life expectancy

```{r ,warning = FALSE, echo=FALSE, message=FALSE}
#install.packages(c("ggplot2"))
gg_hist <- ggplot(data = pop, aes(x = lifeExp)) + 
  # Customize bar color and number
  geom_histogram(fill = "blue", color = "black", bins = 10) + 
  # Add title
  ggtitle("Life expectancy for the gap dataset") + 
  # Change x-axis label
  xlab("Life expectancy (years)") + 
  # y-axis label
  ylab("Frequency") + 
  # Change background theme
  theme_classic() 
print(gg_hist)
```

Boxplot of the life expectancy by continent

```{r ,warning = FALSE, echo=FALSE, message=FALSE}
## Now, let's try boxplots, ggplot style:
gg_box <- ggplot(data = pop, 
                 aes(x = continent, 
                     y = lifeExp, 
                     fill = continent)) + 
  geom_boxplot() + 
  ggtitle("Boxplots for lifeExp by continent") + 
  xlab("Continent") + 
  ylab("Life expectancy (years)") +
  theme_minimal() 
print(gg_box)
```

We can also add a linear model on top of plots rather easily. This connects our visualizations to models that we can build using our data, to better understand the relationships among the features. We can do this using the geom_smooth() function:

```{r ,warning = FALSE, echo=FALSE, message=FALSE}
gg_line <- ggplot(pop, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth(method = "lm", lwd = 1, col = "red") + 
  ggtitle("gdpPercap and lifeExp") + 
  xlab("GDP per capita income (USD)") + 
  ylab("Life expectancy (years)") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 40)) +
  scale_x_log10()
print(gg_line)
```

